Carrier phase and amplitude estimation of $M$-ary phase shift keying using pilots and data

### 2012-05-13 paper.tex
Added to paper.

\begin{theorem}\label{thm:consistency} (Almost sure convergence)
Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symetric complex random variables with $\prob \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be observed noisy symbols given by the signal model~\eqref{eq:sigmod}.   Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.  Put $L = \abs{P \cup D}$ and define real numbers $p$ and $d$ as the limits
\[
\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
\] 
Let $R_i \geq 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
\begin{equation}\label{eq:RiandPhii}
R_ie^{j\Phi_i} = 1 + \frac{1}{\rho_0} w_i e^{-j\theta_0} s_i^*,
\end{equation}
and define the continuous function
\[
G(x) = p h_1(x) + d h_2(x)
\]
where
\[
h_1(x) = \prob R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \prob R_1 \cos\sfracpart{ x + \Phi_1}.
\]
If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
\begin{enumerate}
\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
\end{enumerate}
\end{theorem}

\begin{theorem}\label{thm:normality} (Asymptotic normality)
Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
\[
g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
\]
Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0 G(0)$. If the periodic function $g\fracpart{\phi}_{\pi}$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for every integer $k$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
\[
C = \frac{1}{(p + H d)^2} \left( \begin{array}{cc} 
A_1 + A_2 & ? \\
? & ?
\end{array} \right)
\]
as $L \rightarrow \infty$, where
\[
H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
\]
and
\[
A_1 = \prob R_1^2\sin^2(\Phi_1), \qquad A_2 = \prob R_1^2\sin^2\fracpart{\Phi_1}.
\]
\end{theorem}

@@

We are interested in analysing $\hat{\theta}$, the maximiser of $\Re(Z(\theta))^2$ or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.  Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
\begin{align*}
y_i &= a_0 s_i + w_i \\
&= \rho_0 s_i e^{j \theta_0}\left( 1 + \frac{1}{\rho_0} w_i s_i^* e^{-j \theta_0} \right) \\
&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
\end{align*}
%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta}, we have
\begin{align*}
\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
&= \round{ \lambda + \Phi_i + \angle{s_i} },
\end{align*}
where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, then, when $i \in D$,
\begin{align*}
 y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \\
&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}}
\end{align*}
since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  When $i \in P$,  
\[
y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
\]
Now we can write
\[
Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}
\]
and
\[
\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
\] 
Let 
\[
G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
\]
and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle$.  Then $\hat{\lambda}_L$ is the maximiser of $G_L(\lambda)^2$ and equivalently the maximiser of $\abs{G_L(\lambda)}$.  We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.  From this, the proof of strong consistency follows.

@@

and since $v_k(\lambda,\phi) = 0$ for $\phi \in [0,2\pi)$ when $k \notin \{0,1,\dots,M\}$ the infinite sum above can be truncated so that

@@

k_2(\lambda) = \sum_{k = 0}^M \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi

@@

@@

Now, since $g\fracpart{\phi}_\pi$ is continuous at $(2k+1)\tfrac{\pi}{M}$ for each integer $k$, and $v_k(\phi)$ is continuous and $v_k(\tfrac{(2k+1)\pi}{M}) =  2\sin(\tfrac{\pi}{M})$,

@@

k_2(\lambda) &= \sum_{k = 0}^M \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) + \zeta_k(\lambda) A_k ) \\

@@

k_2(\hat{\lambda}_L) = o_P(1) + 2 \hat{\lambda}_L \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})

>>>>

\begin{theorem}\label{thm:consistency} (Almost sure convergence)
Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symetric complex random variables with $\prob \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be noisy symbols given by the signal model~\eqref{eq:sigmod}.   Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.  Put $L = \abs{P \cup D}$ and define real numbers $p$ and $d$ as the limits
\[
\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
\] 
Let $R_i \geq 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
\begin{equation}\label{eq:RiandPhii}
R_ie^{j\Phi_i} = 1 + \frac{1}{\rho_0} w_i e^{-j\theta_0} s_i^*,
\end{equation}
and define the continuous function
\[
G(x) = p h_1(x) + d h_2(x)
\]
where
\[
h_1(x) = \prob R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \prob R_1 \cos\sfracpart{ x + \Phi_1}.
\]
If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
\begin{enumerate}
\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
\end{enumerate}
\end{theorem}

\begin{theorem}\label{thm:normality} (Asymptotic normality)
Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
\[
g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
\]
Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0 G(0)$. If $g(\phi)$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for every integer $k$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
\[
C = \frac{1}{(p + H d)^2} \left( \begin{array}{cc} 
A_1 + A_2 & ? \\
? & ?
\end{array} \right)
\]
as $L \rightarrow \infty$, where
\[
H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M})
\]
and
\[
A_1 = \prob R_1^2\sin^2(\Phi_1), \qquad A_2 = \prob R_1^2\sin^2\fracpart{\Phi_1}.
\]
\end{theorem}

@@

We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$ or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.  Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
\begin{align*}
y_i &= a_0 s_i + w_i \\
&= \rho_0 s_i e^{j \theta_0}\left( 1 + \frac{1}{\rho_0} w_i s_i^* e^{-j \theta_0} \right) \\
&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
\end{align*}
%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta}, we have
\begin{align*}
\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
&= \round{ \lambda + \Phi_i + \angle{s_i} },
\end{align*}
where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, then, when $i \in D$,
\begin{align*}
 y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \\
&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}}
\end{align*}
since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  When $i \in P$,  
\[
y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
\]
Now we can write
\[
Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}
\]
and
\[
\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
\] 
Let 
\[
G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
\]
and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle$.  Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.  We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.  The proof of part 1 of the theorem follows from this.

@@

and since $v_k(\lambda,\phi) = 0$ for $\phi \in [0,2\pi)$ when $k \notin \{0,1,\dots,M-1\}$ the infinite sum above can be truncated so that

@@

k_2(\lambda) = \sum_{k = 0}^{M-1} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi

@@

(BLERG this needs modification, more care)

@@

Now, since $g(\phi)$ is continuous at $(2k+1)\tfrac{\pi}{M}$ for each integer $k$, and $v_k(\phi)$ is continuous and $v_k(\tfrac{(2k+1)\pi}{M}) =  2\sin(\tfrac{\pi}{M})$,

@@

k_2(\lambda) &= \sum_{k = 0}^{M-1} \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) + \zeta_k(\lambda) A_k ) \\

@@

k_2(\hat{\lambda}_L) = o_P(1) + 2 \hat{\lambda}_L \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M})
